---
title: "Homework 7"
author: "Shelby, Seana, and Greg"
date: "2023-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(classdata)
library(ggplot2)
library(tidyverse)
data(happy, package="productplots")
head(happy)
```

Loading extra data into R

```{r}
HAPPY <- readRDS("data/HAPPY.rds")
str(HAPPY)
```


1.Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA. Show with bar/histogram

```{r, error=TRUE}
str(HAPPY)
HAPPY <- replace(HAPPY, HAPPY == "IAP", NA)
HAPPY <- replace(HAPPY, HAPPY == "DK", NA)
HAPPY <- replace(HAPPY, HAPPY == "NA", NA)
HAPPY <- replace(HAPPY, SEX == "IAP", NA)
HAPPY <- replace(HAPPY, SEX == "DK", NA)
HAPPY <- replace(HAPPY, SEX == "NA", NA)
HAPPY <- replace(HAPPY, MARITAL == "IAP", NA)
HAPPY <- replace(HAPPY, MARITAL == "DK", NA)
HAPPY <- replace(HAPPY, MARITAL == "NA", NA)
HAPPY <- replace(HAPPY, DEGREE == "IAP", NA)
HAPPY <- replace(HAPPY, DEGREE == "DK", NA)
HAPPY <- replace(HAPPY, DEGREE == "NA", NA)
HAPPY <- replace(HAPPY, FINRELA == "IAP", NA)
HAPPY <- replace(HAPPY, FINRELA == "DK", NA)
HAPPY <- replace(HAPPY, FINRELA == "NA", NA)
HAPPY <- replace(HAPPY, HEALTH == "IAP", NA)
HAPPY <- replace(HAPPY, HEALTH == "DK", NA)
HAPPY <- replace(HAPPY, HEALTH == "NA", NA)
HAPPY <- replace(HAPPY, PARTYID == "IAP", NA)
HAPPY <- replace(HAPPY, PARTYID == "DK", NA)
HAPPY <- replace(HAPPY, PARTYID == "NA", NA)
HAPPY <- replace(HAPPY, POLVIEWS == "IAP", NA)
HAPPY <- replace(HAPPY, POLVIEWS == "DK", NA)
HAPPY <- replace(HAPPY, POLVIEWS == "NA", NA)

str(HAPPY)
```

```{r}
ggplot(HAPPY, aes(x = HAPPY)) + geom_bar()
ggplot(HAPPY, aes(x = SEX)) + geom_bar()
ggplot(HAPPY, aes(x = MARITAL)) + geom_bar()
ggplot(HAPPY, aes(x = DEGREE)) + geom_bar()
ggplot(HAPPY, aes(x = FINRELA)) + geom_bar()
ggplot(HAPPY, aes(x = HEALTH)) + geom_bar()
ggplot(HAPPY, aes(x = PARTYID)) + geom_bar()
ggplot(HAPPY, aes(x = POLVIEWS)) + geom_bar()


```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.Show with bar/histogram
```{r}
HAPPY$HAPPY  <- as.factor(HAPPY$HAPPY)
HAPPY$SEX    <- as.factor(HAPPY$SEX)
HAPPY$AGE.num    <- as.numeric(replace(HAPPY$AGE, HAPPY$AGE == "89 OR OLDER", 89))
HAPPY$MARITAL<- as.factor(HAPPY$MARITAL)
HAPPY$DEGREE <- as.factor(HAPPY$DEGREE)
HAPPY$FINRELA<- as.factor(HAPPY$FINRELA)
HAPPY$HEALTH <- as.factor(HAPPY$HEALTH)
HAPPY$PARTYID<- as.factor(HAPPY$PARTYID)
HAPPY$POLVIEWS<- as.factor(HAPPY$POLVIEWS)
```


3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age. Show with bar/histogram

```{r}

HAPPY <- HAPPY %>%
  mutate(DEGREE=fct_relevel(DEGREE, "LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE", "NA's"))

HAPPY <- HAPPY %>%
  mutate(MARITAL=fct_relevel(MARITAL, "MARRIED", "NEVER MARRIED", "SEPARATED", "DIVORCED", "WIDOWED", "NA's"))

HAPPY <- HAPPY %>%
  mutate(FINRELA=fct_relevel(FINRELA, "FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE", "NA's"))

HAPPY <- HAPPY %>%
  mutate(HEALTH=fct_relevel(HEALTH, "POOR", "FAIR", "GOOD", "EXCELLENT", "NA's"))

HAPPY <- HAPPY %>%
  mutate(PARTYID=fct_relevel(PARTYID, "STRONG REPUBLICAN", "NOT STR REPUBLICAN", "IND,NEAR REP", "INDEPENDENT", "IND,NEAR DEM", "NOT STR DEMOCRAT", "STRONG DEMOCRAT", "OTHER PARTY", "NA's"))

HAPPY <- HAPPY %>%
  mutate(POLVIEWS=fct_relevel(POLVIEWS, "EXTRMLY CONSERVATIVE", "CONSERVATIVE", "SLGHTLY CONSERVATIVE", "MODERATE", "SLIGHTLY LIBERAL", "LIBERAL", "EXTREMELY LIBERAL", "NA's"))

ggplot(HAPPY, aes(x = HAPPY)) + geom_bar()
ggplot(HAPPY, aes(x = SEX)) + geom_bar()
ggplot(HAPPY, aes(x = MARITAL)) + geom_bar()
ggplot(HAPPY, aes(x = DEGREE)) + geom_bar()
ggplot(HAPPY, aes(x = FINRELA)) + geom_bar()
ggplot(HAPPY, aes(x = HEALTH)) + geom_bar()
ggplot(HAPPY, aes(x = PARTYID)) + geom_bar()
ggplot(HAPPY, aes(x = POLVIEWS)) + geom_bar()

```

4.Investigate the relationship between happiness and two other variables in the data. Find a visualization that captures the relationship and write a paragraph to describe it.

Each member should study a different question (choose different variables to investigate).

Upload your answers and the dataset to the repository. Discuss and refine answers as a team.

Shelby-
Seana-What is the relationship between happiness and political views? Happiness and Degree?
For political views, happiness appears to be slightly higher in extremely conservative people, and is about even between males and females. Happiness appears to be lowest in extremely liberal people and is slightly lower in males than females for that level. Happiness is highest in those with graduate degrees and is slightly higher in females

```{r}

HAPPY$m.happy <- as.numeric(HAPPY$HAPPY, na.rm=TRUE)

ggplot(HAPPY, aes(x = m.happy, fill = POLVIEWS)) + geom_bar()

ggplot(HAPPY, aes(x = m.happy, fill = DEGREE)) + geom_bar()



HAPPY %>%
  ggplot(aes(fill = SEX, y=m.happy, x=POLVIEWS)) +
  geom_bar(position="dodge", stat="summary") + 
  ylim(0,5)

HAPPY %>%
  ggplot(aes(fill = SEX, y=m.happy, x=DEGREE)) +
  geom_bar(position="dodge", stat="summary") + 
  ylim(0,5)
```

Greg- HAPPY, AGE, DEGREE
```{r}
HAPPY %>% ggplot(aes(x=HAPPY, weight=AGE.num, fill=DEGREE)) + geom_bar()
```
  
